Derivation of Variational Bayes

### 2016-09-22 AR2again.tex
updates

\begin{document}
\SweaveOpts{concordance=TRUE}
%\SweaveOpts{concordance=TRUE}

@@

\begin{figure}
<<MCMC>>=
setwd("/home/nltom2/Desktop/Confirmation")
library(ggplot2)
library(GGally)
thin <- read.csv("AR2thin.csv")[,-1]
ggpairs(thin)
@
\caption{Result of MCMC with 120,000 draws. The first 20\% were discarded then the remaining draws were thinned by a factor of 20, which was required to remove any dependence. The remaining effective sample size for each parameter is 5000.
\end{figure}

<<summary>>=
library(knitr)
library(mvtnorm)
library(reshape)
summary = matrix( c(0.78, 0.16, 0.10, -0.09, 0.05, 48.9, 77.3, -76.7, 2.6,
                    0.83, 0.10, 0.04, 0, 0.05, 50, 79.7, -77.7, "1.3 x 10^{-6}", 
                    0.63, 0.27, 0.09, -0.12, 0.05, 45.6, 81.4, -78.3, 51.1, 
                    0.78, 0.15, 0.1, -0.09, 0.05, 49.3, 78, -76.7, 4.0,
                    0.77, 0.15, 0.09, -0.09, 0.05, 48.5, 77.6, -76.9, 1.3) , byrow = TRUE, ncol = 9)

colnames(summary) = c(paste0(expression(mu), 1), paste0(expression(mu), 2), "L11", "L21", "L22", expression(alpha), expression(beta), "L(theta)", "time (seconds)*")
rownames(summary) = c("Method of Moments", "Mean Field", "SVB", "SVB CV", "SVB RP")
kable(summary)
@
The original stochastic variational bayes algorithm converged to a different local maximum, possibly due to the high variance in its gradient estimates.

>>>>

\usepackage{Sweave}
\begin{document}
\input{AR2again-concordance}
%\SweaveOpts{concordance=TRUE}

@@

\begin{figure}
\begin{Schunk}
\begin{Sinput}
> setwd("/home/nltom2/Desktop/Confirmation")
> library(ggplot2)
> library(GGally)
> thin <- read.csv("AR2thin.csv")[,-1]
> ggpairs(thin)
\end{Sinput}
\end{Schunk}
\caption{Result of MCMC with 120,000 draws. The first 20\% were discarded then the remaining draws were thinned by a factor of 20, which was required to remove any dependence. The remaining effective sample size for each parameter is 5000.
\end{figure}

\begin{Schunk}
\begin{Sinput}
> library(knitr)
> library(mvtnorm)
> library(reshape)
> summary = matrix( c(0.78, 0.16, 0.10, -0.09, 0.05, 48.9, 77.3, -76.7, 2.6,
+                     0.83, 0.10, 0.04, 0, 0.05, 50, 79.7, -77.7, "1.3 x 10^{-6}", 
+                     0.63, 0.27, 0.09, -0.12, 0.05, 45.6, 81.4, -78.3, 51.1, 
+                     0.78, 0.15, 0.1, -0.09, 0.05, 49.3, 78, -76.7, 4.0,
+                     0.77, 0.15, 0.09, -0.09, 0.05, 48.5, 77.6, -76.9, 1.3) , byrow = TRUE, ncol = 9)
> colnames(summary) = c(paste0(expression(mu), 1), paste0(expression(mu), 2), "L11", "L21", "L22", expression(alpha), expression(beta), "L(theta)", "time (seconds)*")
> rownames(summary) = c("Method of Moments", "Mean Field", "SVB", "SVB CV", "SVB RP")
> kable(summary)
\end{Sinput}
\begin{Soutput}
|                  |mu1  |mu2  |L11  |L21   |L22  |alpha |beta |L(theta) |time (seconds)* |
|:-----------------|:----|:----|:----|:-----|:----|:-----|:----|:--------|:---------------|
|Method of Moments |0.78 |0.16 |0.1  |-0.09 |0.05 |48.9  |77.3 |-76.7    |2.6             |
|Mean Field        |0.83 |0.1  |0.04 |0     |0.05 |50    |79.7 |-77.7    |1.3 x 10^{-6}   |
|SVB               |0.63 |0.27 |0.09 |-0.12 |0.05 |45.6  |81.4 |-78.3    |51.1            |
|SVB CV            |0.78 |0.15 |0.1  |-0.09 |0.05 |49.3  |78   |-76.7    |4               |
|SVB RP            |0.77 |0.15 |0.09 |-0.09 |0.05 |48.5  |77.6 |-76.9    |1.3             |
\end{Soutput}
\end{Schunk}
The original stochastic variational bayes algorithm converged to a different local maximum, possibly due to the high variance in its gradient estimates.

### 2016-09-22 written report/2levelmixture.tex
updates

\documentclass[12pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[colorlinks=TRUE, linkcolor=blue]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage[aboveskip=2pt]{subcaption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{ulem}
\usepackage{afterpage}
\usepackage{bbm}

\graphicspath{{images/}}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

The quantity $\tilde{\pi}_1 = \mathbb{E}_\pi \ln(\pi) = \psi(\alpha)-\psi(\alpha+\beta)$, and $\tilde{\pi}_2 = \mathbb{E}_\pi \ln(1-\pi) = \psi(\beta)-\psi(\alpha+\beta)$, where $\psi(\cdot)$ is the digamma function (Archambeau and Verleysen 2007). We also expanded the quadratic term and substitued in $(y_i^2 - \bar{\mu}_j)^2 + \lambda_j$ for $\mathbb{E} (y_i-\mu_j)^2$.

@@

\begin{eqnarray*}
a & = & \sum_{i=1}^{n} \frac{p_i}{p_i+q_i} + 1 \\
b & = & \sum_{i=1}^{n} \frac{q_i}{p_i+q_i} + 1 \\
\bar{\mu}_1 & = &  \frac{\sum_{i=1}^{n} y_i p_i /(p_i+q_i)}{\sum_{i=1}^{n} p_i /(p_i+q_i)} \\
\lambda_1 & = & \left(\sum_{i=1}^{n} \frac{p_i}{p_i+q_i}\right)^{-1} \\
\bar{\mu}_2 & = &  \frac{\sum_{i=1}^{n} y_i q_i /(p_i+q_i)}{\sum_{i=1}^{n} q_i /(p_i+q_i)} \\
\lambda_2 & = & \left(\sum_{i=1}^{n} \frac{q_i}{p_i+q_i}\right)^{-1} \\
p_i & = & \exp \left\{\frac{2(\psi(a)-\psi(a+b))-((y_i - \bar{\mu}_1)^2 + \lambda_1)}{2} \right\} \\
q_i & = & \exp \left\{\frac{2(\psi(b)-\psi(a+b))-((y_i - \bar{\mu}_2)^2 + \lambda_2)}{2} \right\}
\end{eqnarray*}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{figure/vbayes-1}

\end{knitrout}
250 draws were simulated with parameters $\mu_1 = 3, \mu_2 = 6, \pi=0.6$ and th e variational algorithm was ran. After manually correcting mislabeling, $y_i$ was allocated to distribution 1 if $p_i > q_i$ and to distribution 2 if $p_i < q_i$, resulting in the successful classification of 235/250 draws. A more trivial decision rule to allocate $y_i$ to distribution 1 if $y_i < \bar{y}$ and to distribution 2 if $y_i > \bar{y}$ successfully classified 234/250 draws.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
\includegraphics[width=\maxwidth]{figure/vbayes2-1}

\end{knitrout}
250 draws were simulated with parameters $\mu_1 = 5.5, \mu_2 = 6, \pi=0.6$ to try and force an overlap in the data. $y_i$ was allocated to distribution 1 if $p_i > q_i$ and to distribution 2 if $p_i < q_i$, resulting in the successful classification of 158/250 draws. The trivial decision rule had identical classifications.

>>>>

\documentclass[12pt]{article}
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[colorlinks=TRUE, linkcolor=blue]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage[aboveskip=2pt]{subcaption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{ulem}
\usepackage{afterpage}
\usepackage{bbm}

\graphicspath{{images/}}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\usepackage{Sweave}
\begin{document}
\input{2levelmixture-concordance}

@@

The quantity $\tilde{\pi}_1 = \mathbb{E}_\pi \ln(\pi) = \psi(\alpha)-\psi(\alpha+\beta)$, and $\tilde{\pi}_2 = \mathbb{E}_\pi \ln(1-\pi) = \psi(\beta)-\psi(\alpha+\beta)$, where $\psi(\cdot)$ is the digamma function (Archambeau and Verleysen 2007).

@@

\begin{eqnarray*}
\alpha & = & \sum_{i=1}^{n} \frac{p_i}{p_i+q_i} + 1 \\
\beta & = & \sum_{i=1}^{n} \frac{q_i}{p_i+q_i} + 1 \\
\bar{\mu}_1 & = &  \frac{\sum_{i=1}^{n} y_i p_i /(p_i+q_i)}{\sum_{i=1}^{n} p_i /(p_i+q_i)} \\
\lambda_1 & = & \left(\sum_{i=1}^{n} \frac{p_i}{p_i+q_i}\right)^{-1} \\
\bar{\mu}_2 & = &  \frac{\sum_{i=1}^{n} y_i q_i /(p_i+q_i)}{\sum_{i=1}^{n} q_i /(p_i+q_i)} \\
\lambda_2 & = & \left(\sum_{i=1}^{n} \frac{q_i}{p_i+q_i}\right)^{-1} \\
p_i & = & \exp \left\{\frac{2(\psi(\alpha)-\psi(\alpha+\beta))-((y_i - \bar{\mu}_1)^2 + \lambda_1)}{2} \right\} \\
q_i & = & \exp \left\{\frac{2(\psi(\beta)-\psi(\alpha+\beta))-((y_i - \bar{\mu}_2)^2 + \lambda_2)}{2} \right\}
\end{eqnarray*}

250 draws were simulated with parameters $\mu_1 = 3, \mu_2 = 6, \pi=0.6$ and th e variational algorithm was ran. After manually correcting mislabeling, $y_i$ was allocated to distribution 1 if $p_i > q_i$ and to distribution 2 if $p_i < q_i$, resulting in the successful classification of 235/250 draws. A more trivial decision rule to allocate $y_i$ to distribution 1 if $y_i < \bar{y}$ and to distribution 2 if $y_i > \bar{y}$ successfully classified 234/250 draws.

250 draws were simulated with parameters $\mu_1 = 5.5, \mu_2 = 6, \pi=0.6$ to try and force an overlap in the data. $y_i$ was allocated to distribution 1 if $p_i > q_i$ and to distribution 2 if $p_i < q_i$, resulting in the successful classification of 158/250 draws. The trivial decision rule had identical classifications.

### 2016-09-22 written report/AR2.tex
updates

\begin{document}
\SweaveOpts{concordance=TRUE}
\section{The AR2 Model}

@@

<<summary, Echo = FALSE, Eval = TRUE>>=
library(knitr)
library(mvtnorm)
library(reshape)
summary = matrix( c(0.37, -0.30, 0.10, -0.03, 0.10, 49.2, 78.4, -77.5, 7.4,
                    0.37, -0.29, 0.09, 0, 0.10, 49.5, 79.8, -77.4, "1.3 x 10^{-6}", 
                    0.32, -0.32, 0.11, -0.01, 0.12, 46.4, 78.7, -77.2, 59.5, 
                    0.38, -0.30, -0.1, 0.03, -0.10, 48.7, 78.9, -77.4, 1.2) , byrow = TRUE, ncol = 9)

colnames(summary) = c(paste0(expression(mu), 1), paste0(expression(mu), 2), "L11", "L21", "L22", expression(alpha), expression(beta), "L(theta)", "time (seconds)*")
rownames(summary) = c("Method of Moments", "Mean Field", "Copula 1", "Copula 2")
kable(summary)
@
\*Note that all of the variational algorithms are currently unparallelised, and the copula versions in particular have a huge potential for faster performance.

@@

\begin{figure}
<<MCMC, Echo = FALSE>>=
setwd("/home/nltom2/Desktop/Code/Approximate-Updating")
library(ggplot2)
library(GGally)
thin <- read.csv("AR2thin.csv")[,-1]
ggpairs(thin)
@
\end{figure}

@@

<<copula, Echo = FALSE>>=

copest =data.frame(copula = c("gaussian", "t"), rho = c(-0.29, -0.29), df = c("NA", 112.5), BIC = c(-167.8, -161.2))
kable(copest)
@

@@

<<MM, Echo = FALSE>>=
MM = matrix( c(0.37, -0.30, 0.10, -0.03, 0.10, 49.2, 78.4, -77.5, 7.4), ncol = 9)
colnames(MM) = c(paste0(expression(mu), 1), paste0(expression(mu), 2), "L11", "L21", "L22", expression(alpha), expression(beta), "L(theta)", "time (seconds)")
kable(MM)
@

@@

<<MFSummary, Echo = FALSE>>=
MFsum = matrix(c(0.37, -0.29, 0.09, 0, 0.10, 49.5, 79.8, -77.4, "1.3 x 10^{-6}"), ncol = 9)
colnames(MFsum) = c(paste0(expression(mu), 1), paste0(expression(mu), 2), "L11", "L21", "L22", expression(alpha), expression(beta), "L(theta)", "time (seconds)")
kable(MFsum)
@

\begin{figure}
<<meanfield, Echo = FALSE>>=
library(pscl)
library(gridExtra)
mf <- read.csv("AR2mf.csv")[,-1]
astar = 50
xp1 = seq(-0.1, 1, length.out=1000)
p1dens = dnorm(xp1, mf[10, 2], sqrt(mf[10, 3]))
p1dat = data.frame(cbind(xp1, p1dens))
p1plot = ggplot(p1dat, aes(x=xp1, y=p1dens)) + geom_line() + labs(x="Phi 1", y=NULL)

xp2 = seq(-0.8, 0.4, length.out=1000)
p2dens = dnorm(xp2, mf[10, 4], sqrt(mf[10, 5]))
p2dat = data.frame(cbind(xp2, p2dens))
p2plot = ggplot(p2dat, aes(x=xp2, y=p2dens)) + geom_line() + labs(x="Phi 2", y=NULL)

xs = seq(0.5, 3.5, length.out = 1000)
sdens = densigamma(xs, astar, mf[10, 1])
sigdat = data.frame(cbind(xs, sdens))
sigplot = ggplot(sigdat, aes(x=xs, y=sdens)) + geom_line() + labs(x="Sigma Squared", y=NULL)

grid.arrange(p1plot, p2plot, sigplot, ncol = 2)
@
\end{figure}

@@

\begin{figure}
<<SGA1, Echo = FALSE>>=
SGA1 = matrix( c(0.32, -0.32, 0.11, -0.01, 0.12, 46.4, 78.7, -77.2, 59.5), ncol = 9)
colnames(SGA1) = c(paste0(expression(mu), 1), paste0(expression(mu), 2), "L11", "L21", "L22", expression(alpha), expression(beta), "L(theta)", "time (seconds)")
kable(SGA1)
L2 =  c(0.32, -0.32, 0.11, -0.01, 0.12, 46.4, 78.7)
L = matrix(c(L2[3:4], 0, L2[5]), 2)
Sigma = L %*% t(L)

xp1 = seq(-0.1, 1, length.out=1000)
p1dens = dnorm(xp1, L2[1], sqrt(Sigma[1, 1]))
p1dat = data.frame(cbind(xp1, p1dens))
p1plot = ggplot(p1dat, aes(x=xp1, y=p1dens)) + geom_line() + labs(x="Phi 1", y=NULL)

xp2 = seq(-0.8, 0.4, length.out=1000)
p2dens = dnorm(xp2, L2[2], sqrt(Sigma[2, 2]))
p2dat = data.frame(cbind(xp2, p2dens))
p2plot = ggplot(p2dat, aes(x=xp2, y=p2dens)) + geom_line() + labs(x="Phi 2", y=NULL)

xs = seq(0.5, 3.5, length.out = 1000)
sdens = densigamma(xs, L2[6], L2[7])
sigdat = data.frame(cbind(xs, sdens))
sigplot = ggplot(sigdat, aes(x=xs, y=sdens)) + geom_line() + labs(x="Sigma Squared", y=NULL)

x1 = seq(-1, 1, length.out=1000)
x2 = seq(-1, 1, length.out=1000)
z = matrix(0, length(x1), length(x2))
for (i in 1:length(x1)) {
  d1 = x1
  d2= x2[i]
  z[,i] = dmvnorm(cbind(d1,d2), L2[1:2], Sigma)
}
colnames(z) = x1
rownames(z) = x2
# reshape the data
dat.p1p2 <- melt(z)
colnames(dat.p1p2) = c("Phi1", "Phi2", "value")

p1p2 <- ggplot(dat.p1p2, aes(x=Phi1, y=Phi2, z = value)) + geom_contour()+ coord_equal()
grid.arrange(p1plot, p2plot, sigplot, p1p2)
@
\end{figure}
\section{Copula Variational Bayes 2}

@@

\begin{figure}
<<SGA2, Echo = FALSE>>=
SGA2 = matrix( c(0.38, -0.30, -0.1, 0.03, -0.10, 48.7, 78.9, -77.4, 1.2), ncol = 9)
colnames(SGA2) = c(paste0(expression(mu), 1), paste0(expression(mu), 2), "L11", "L21", "L22", expression(alpha), expression(beta), "L(theta)", "time (seconds)")
kable(SGA2)
L2 =  c(0.38, -0.30, -0.1, 0.03, -0.10, 48.7, 78.9)
L = matrix(c(L2[3:4], 0, L2[5]), 2)
Sigma = L %*% t(L)

xp1 = seq(-0.1, 1, length.out=1000)
p1dens = dnorm(xp1, L2[1], sqrt(Sigma[1, 1]))
p1dat = data.frame(cbind(xp1, p1dens))
p1plot = ggplot(p1dat, aes(x=xp1, y=p1dens)) + geom_line() + labs(x="Phi 1", y=NULL)

xp2 = seq(-0.8, 0.4, length.out=1000)
p2dens = dnorm(xp2, L2[2], sqrt(Sigma[2, 2]))
p2dat = data.frame(cbind(xp2, p2dens))
p2plot = ggplot(p2dat, aes(x=xp2, y=p2dens)) + geom_line() + labs(x="Phi 2", y=NULL)

xs = seq(0.5, 3.5, length.out = 1000)
sdens = densigamma(xs, L2[6], L2[7])
sigdat = data.frame(cbind(xs, sdens))
sigplot = ggplot(sigdat, aes(x=xs, y=sdens)) + geom_line() + labs(x="Sigma Squared", y=NULL)

x1 = seq(-1, 1, length.out=1000)
x2 = seq(-1, 1, length.out=1000)
z = matrix(0, length(x1), length(x2))
for (i in 1:length(x1)) {
  d1 = x1
  d2= x2[i]
  z[,i] = dmvnorm(cbind(d1,d2), L2[1:2], Sigma)
}
colnames(z) = x1
rownames(z) = x2
# reshape the data
dat.p1p2 <- melt(z)
colnames(dat.p1p2) = c("Phi1", "Phi2", "value")

p1p2 <- ggplot(dat.p1p2, aes(x=Phi1, y=Phi2, z = value)) + geom_contour()+ coord_equal()
grid.arrange(p1plot, p2plot, sigplot, p1p2)
@
\end{figure}
\end{document}

>>>>

\usepackage{Sweave}
\begin{document}
\input{AR2-concordance}
\section{The AR2 Model}

@@

\begin{Schunk}
\begin{Sinput}
> library(knitr)
> library(mvtnorm)
> library(reshape)
> summary = matrix( c(0.37, -0.30, 0.10, -0.03, 0.10, 49.2, 78.4, -77.5, 7.4,
+                     0.37, -0.29, 0.09, 0, 0.10, 49.5, 79.8, -77.4, "1.3 x 10^{-6}", 
+                     0.32, -0.32, 0.11, -0.01, 0.12, 46.4, 78.7, -77.2, 59.5, 
+                     0.38, -0.30, -0.1, 0.03, -0.10, 48.7, 78.9, -77.4, 1.2) , byrow = TRUE, ncol = 9)
> colnames(summary) = c(paste0(expression(mu), 1), paste0(expression(mu), 2), "L11", "L21", "L22", expression(alpha), expression(beta), "L(theta)", "time (seconds)*")
> rownames(summary) = c("Method of Moments", "Mean Field", "Copula 1", "Copula 2")
> kable(summary)
\end{Sinput}
\begin{Soutput}
|                  |mu1  |mu2   |L11  |L21   |L22  |alpha |beta |L(theta) |time (seconds)* |
|:-----------------|:----|:-----|:----|:-----|:----|:-----|:----|:--------|:---------------|
|Method of Moments |0.37 |-0.3  |0.1  |-0.03 |0.1  |49.2  |78.4 |-77.5    |7.4             |
|Mean Field        |0.37 |-0.29 |0.09 |0     |0.1  |49.5  |79.8 |-77.4    |1.3 x 10^{-6}   |
|Copula 1          |0.32 |-0.32 |0.11 |-0.01 |0.12 |46.4  |78.7 |-77.2    |59.5            |
|Copula 2          |0.38 |-0.3  |-0.1 |0.03  |-0.1 |48.7  |78.9 |-77.4    |1.2             |
\end{Soutput}
\end{Schunk}
\*Note that all of the variational algorithms are currently unparallelised, and the copula versions in particular have a huge potential for faster performance.

@@

\begin{figure}
\begin{Schunk}
\begin{Sinput}
> setwd("/home/nltom2/Desktop/Code/Approximate-Updating")
> library(ggplot2)
> library(GGally)
> thin <- read.csv("AR2thin.csv")[,-1]
> ggpairs(thin)
\end{Sinput}
\end{Schunk}
\end{figure}

@@

\begin{Schunk}
\begin{Sinput}
> copest =data.frame(copula = c("gaussian", "t"), rho = c(-0.29, -0.29), df = c("NA", 112.5), BIC = c(-167.8, -161.2))
> kable(copest)
\end{Sinput}
\begin{Soutput}
|copula   |   rho|df    |    BIC|
|:--------|-----:|:-----|------:|
|gaussian | -0.29|NA    | -167.8|
|t        | -0.29|112.5 | -161.2|
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> MM = matrix( c(0.37, -0.30, 0.10, -0.03, 0.10, 49.2, 78.4, -77.5, 7.4), ncol = 9)
> colnames(MM) = c(paste0(expression(mu), 1), paste0(expression(mu), 2), "L11", "L21", "L22", expression(alpha), expression(beta), "L(theta)", "time (seconds)")
> kable(MM)
\end{Sinput}
\begin{Soutput}
|  mu1|  mu2| L11|   L21| L22| alpha| beta| L(theta)| time (seconds)|
|----:|----:|---:|-----:|---:|-----:|----:|--------:|--------------:|
| 0.37| -0.3| 0.1| -0.03| 0.1|  49.2| 78.4|    -77.5|            7.4|
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> MFsum = matrix(c(0.37, -0.29, 0.09, 0, 0.10, 49.5, 79.8, -77.4, "1.3 x 10^{-6}"), ncol = 9)
> colnames(MFsum) = c(paste0(expression(mu), 1), paste0(expression(mu), 2), "L11", "L21", "L22", expression(alpha), expression(beta), "L(theta)", "time (seconds)")
> kable(MFsum)
\end{Sinput}
\begin{Soutput}
|mu1  |mu2   |L11  |L21 |L22 |alpha |beta |L(theta) |time (seconds) |
|:----|:-----|:----|:---|:---|:-----|:----|:--------|:--------------|
|0.37 |-0.29 |0.09 |0   |0.1 |49.5  |79.8 |-77.4    |1.3 x 10^{-6}  |
\end{Soutput}
\end{Schunk}

\begin{figure}
\begin{Schunk}
\begin{Sinput}
> library(pscl)
> library(gridExtra)
> mf <- read.csv("AR2mf.csv")[,-1]
> astar = 50
> xp1 = seq(-0.1, 1, length.out=1000)
> p1dens = dnorm(xp1, mf[10, 2], sqrt(mf[10, 3]))
> p1dat = data.frame(cbind(xp1, p1dens))
> p1plot = ggplot(p1dat, aes(x=xp1, y=p1dens)) + geom_line() + labs(x="Phi 1", y=NULL)
> xp2 = seq(-0.8, 0.4, length.out=1000)
> p2dens = dnorm(xp2, mf[10, 4], sqrt(mf[10, 5]))
> p2dat = data.frame(cbind(xp2, p2dens))
> p2plot = ggplot(p2dat, aes(x=xp2, y=p2dens)) + geom_line() + labs(x="Phi 2", y=NULL)
> xs = seq(0.5, 3.5, length.out = 1000)
> sdens = densigamma(xs, astar, mf[10, 1])
> sigdat = data.frame(cbind(xs, sdens))
> sigplot = ggplot(sigdat, aes(x=xs, y=sdens)) + geom_line() + labs(x="Sigma Squared", y=NULL)
> grid.arrange(p1plot, p2plot, sigplot, ncol = 2)
\end{Sinput}
\end{Schunk}
\end{figure}

@@

\begin{figure}
\begin{Schunk}
\begin{Sinput}
> SGA1 = matrix( c(0.32, -0.32, 0.11, -0.01, 0.12, 46.4, 78.7, -77.2, 59.5), ncol = 9)
> colnames(SGA1) = c(paste0(expression(mu), 1), paste0(expression(mu), 2), "L11", "L21", "L22", expression(alpha), expression(beta), "L(theta)", "time (seconds)")
> kable(SGA1)
\end{Sinput}
\begin{Soutput}
|  mu1|   mu2|  L11|   L21|  L22| alpha| beta| L(theta)| time (seconds)|
|----:|-----:|----:|-----:|----:|-----:|----:|--------:|--------------:|
| 0.32| -0.32| 0.11| -0.01| 0.12|  46.4| 78.7|    -77.2|           59.5|
\end{Soutput}
\begin{Sinput}
> L2 =  c(0.32, -0.32, 0.11, -0.01, 0.12, 46.4, 78.7)
> L = matrix(c(L2[3:4], 0, L2[5]), 2)
> Sigma = L %*% t(L)
> xp1 = seq(-0.1, 1, length.out=1000)
> p1dens = dnorm(xp1, L2[1], sqrt(Sigma[1, 1]))
> p1dat = data.frame(cbind(xp1, p1dens))
> p1plot = ggplot(p1dat, aes(x=xp1, y=p1dens)) + geom_line() + labs(x="Phi 1", y=NULL)
> xp2 = seq(-0.8, 0.4, length.out=1000)
> p2dens = dnorm(xp2, L2[2], sqrt(Sigma[2, 2]))
> p2dat = data.frame(cbind(xp2, p2dens))
> p2plot = ggplot(p2dat, aes(x=xp2, y=p2dens)) + geom_line() + labs(x="Phi 2", y=NULL)
> xs = seq(0.5, 3.5, length.out = 1000)
> sdens = densigamma(xs, L2[6], L2[7])
> sigdat = data.frame(cbind(xs, sdens))
> sigplot = ggplot(sigdat, aes(x=xs, y=sdens)) + geom_line() + labs(x="Sigma Squared", y=NULL)
> x1 = seq(-1, 1, length.out=1000)
> x2 = seq(-1, 1, length.out=1000)
> z = matrix(0, length(x1), length(x2))
> for (i in 1:length(x1)) {
+   d1 = x1
+   d2= x2[i]
+   z[,i] = dmvnorm(cbind(d1,d2), L2[1:2], Sigma)
+ }
> colnames(z) = x1
> rownames(z) = x2
> # reshape the data
> dat.p1p2 <- melt(z)
> colnames(dat.p1p2) = c("Phi1", "Phi2", "value")
> p1p2 <- ggplot(dat.p1p2, aes(x=Phi1, y=Phi2, z = value)) + geom_contour()+ coord_equal()
> grid.arrange(p1plot, p2plot, sigplot, p1p2)
\end{Sinput}
\end{Schunk}
\end{figure}
\section{Copula Variational Bayes 2}

@@

\begin{figure}
\begin{Schunk}
\begin{Sinput}
> SGA2 = matrix( c(0.38, -0.30, -0.1, 0.03, -0.10, 48.7, 78.9, -77.4, 1.2), ncol = 9)
> colnames(SGA2) = c(paste0(expression(mu), 1), paste0(expression(mu), 2), "L11", "L21", "L22", expression(alpha), expression(beta), "L(theta)", "time (seconds)")
> kable(SGA2)
\end{Sinput}
\begin{Soutput}
|  mu1|  mu2|  L11|  L21|  L22| alpha| beta| L(theta)| time (seconds)|
|----:|----:|----:|----:|----:|-----:|----:|--------:|--------------:|
| 0.38| -0.3| -0.1| 0.03| -0.1|  48.7| 78.9|    -77.4|            1.2|
\end{Soutput}
\begin{Sinput}
> L2 =  c(0.38, -0.30, -0.1, 0.03, -0.10, 48.7, 78.9)
> L = matrix(c(L2[3:4], 0, L2[5]), 2)
> Sigma = L %*% t(L)
> xp1 = seq(-0.1, 1, length.out=1000)
> p1dens = dnorm(xp1, L2[1], sqrt(Sigma[1, 1]))
> p1dat = data.frame(cbind(xp1, p1dens))
> p1plot = ggplot(p1dat, aes(x=xp1, y=p1dens)) + geom_line() + labs(x="Phi 1", y=NULL)
> xp2 = seq(-0.8, 0.4, length.out=1000)
> p2dens = dnorm(xp2, L2[2], sqrt(Sigma[2, 2]))
> p2dat = data.frame(cbind(xp2, p2dens))
> p2plot = ggplot(p2dat, aes(x=xp2, y=p2dens)) + geom_line() + labs(x="Phi 2", y=NULL)
> xs = seq(0.5, 3.5, length.out = 1000)
> sdens = densigamma(xs, L2[6], L2[7])
> sigdat = data.frame(cbind(xs, sdens))
> sigplot = ggplot(sigdat, aes(x=xs, y=sdens)) + geom_line() + labs(x="Sigma Squared", y=NULL)
> x1 = seq(-1, 1, length.out=1000)
> x2 = seq(-1, 1, length.out=1000)
> z = matrix(0, length(x1), length(x2))
> for (i in 1:length(x1)) {
+   d1 = x1
+   d2= x2[i]
+   z[,i] = dmvnorm(cbind(d1,d2), L2[1:2], Sigma)
+ }
> colnames(z) = x1
> rownames(z) = x2
> # reshape the data
> dat.p1p2 <- melt(z)
> colnames(dat.p1p2) = c("Phi1", "Phi2", "value")
> p1p2 <- ggplot(dat.p1p2, aes(x=Phi1, y=Phi2, z = value)) + geom_contour()+ coord_equal()
> grid.arrange(p1plot, p2plot, sigplot, p1p2)
\end{Sinput}
\end{Schunk}
\end{figure}
\end{document}

### 2016-09-22 written report/ar1_predictive.Rnw
updates

<<forecast, echo=FALSE>>=
##One step variational forecast
library(Rcpp)
library(tidyr)
library(ggplot2)
library(dplyr)
library(RcppArmadillo)
sourceCpp("AR1c.cpp")
library(truncnorm)
library(flexmix)

>>>>

<<forecast, echo=FALSE>>=
##One step variational forecast
library(Rcpp)
library(tidyr)
library(ggplot2)
library(dplyr)
library(RcppArmadillo)
#sourceCpp("AR1c.cpp")
library(truncnorm)
library(flexmix)
